---
title: "Lab Name"
author: "Author Name"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(openintro)
library(tidycensus)
library(redist)
library(sf)
library(alarmdata)
library(ggnewscale)
library(redistmetrics)
library(dplyr)
```

### Exercise 1
census_api_key("d84f6074a7e95b068d1dbc2a2daeb05bf6ab224b", install = TRUE)



```{r code-chunk-label}


mo_tracts <- get_decennial(
  geography = "tract",
  variables = "P1_001N",
  state = "MO",
  year = 2020,
  geometry = TRUE
) |>
  rename(pop = value)

mo_map <- redist_map(mo_tracts, ndists = 8, pop_tol= 0.001, total_pop = pop)


plot(mo_map)
```

### Exercise 2

```{r}
mo_plans <- redist_smc(mo_map, nsims = 5)
print(mo_plans)

redist.plot.plans(mo_plans, draws=c("1","2","3"), shp= mo_map)
```

```{r}
MO_data <- get_decennial(
  geography = "tract",
  variables = c(
    pop = "P1_001N",
    vap = "P3_001N",
    bvap = "P3_006N",
    hvap = "P3_009N"
  ),
  state = "MO",
  year = 2020,
  geometry = TRUE,
  output = "wide"
)

MO_data <- MO_data |>
  mutate(country = substr(GEOID,1,5))

MO_Minotry_Map <- redist_map(MO_data,
                             ndists = 8,
                             pop_tol = 0.01,
                             total_pop = pop)
  
mo_constr <- redist_constr(MO_Minotry_Map)

# 2. Add Compactness (The simple way)
mo_constr <- add_constr_edges_rem(mo_constr, strength = 0.5)

# 3. Add Minority Power (The correct argument is 'target')
mo_constr <- add_constr_grp_pow(
  mo_constr, 
  group_pop = bvap,
  total_pop = vap,
  tgt_group = 0.45,  
  strength = 10
)

# 4. Run it
mo_MINORplans <- redist_smc(mo_map, nsims = 1000, constraints = mo_constr)

redist.plot.plans(mo_MINORplans, draws = 5,shp =MO_Minotry_Map, qty = bvap / vap)
```
MO_Minor_Plans <- redist_smc(
  map = mo_map,
  nsims = 100,
  counties = county,
  constraints = list(
    grp_pow = list(strenght = 10,
                   group_pop = bvap,
                   total_pop = vap,
                   target_group_prop = 0.45)
  )
)



mo_plans_df <- as_tibble(mo_MINORplans)

# 2. Add your data columns directly
# Since 'precinct' is the row index in your map, this matches them up perfectly
mo_plans_df$pop  <- MO_Minotry_Map$pop[mo_plans_df$precinct]
mo_plans_df$vap  <- MO_Minotry_Map$vap[mo_plans_df$precinct]
mo_plans_df$bvap <- MO_Minotry_Map$bvap[mo_plans_df$precinct]

plan_stats <- mo_MINORplans |>
  group_by(draw,district) |>
  summarize(
    total_pop = sum(pop),
    total_vap = sum(vap),
    total_bvap = sum(bvap),
    .groups = "drop"
  ) |>
  mutate(bvap_pct = total_bvap / total_vap)
  
head(plan_stats)
  
best_vra_plans <- mo_MINORplans |>
  mutate(bvap_pct = bvap_districts) |>
  group_by(draw) |>
  summarize(max_bvap = max(bvap_pct)) |>
  filter(max_bvap > 0.40) |>
  arrange(desc(max_bvap))

print(best_vra_plans)






```{r}
MO_Alarm <- alarm_50state_map("MO")

MO5 <- MO_Alarm |>
  filter( cd_2020 == "5")

MO5_minority <- MO5 |>
  mutate(minority_pct = (vap - vap_white) / vap * 100)

ggplot(MO5_minority) +
  geom_sf(aes(fill = minority_pct), color = NA) +
  scale_fill_viridis_c(name = "% Minority") +
  labs(
    title = "Missouri Congressional District 5",
    subtitle = "Minority Voting Age Population (%)"
  ) +
  theme_minimal()

```

```{r}
MO5_voting <- MO5 |>
  mutate(lean = (adv_20 - arv_20)/ (adv_20 + arv_20))

MO5_voting <- MO5_voting |>
  mutate(
    dem_lean = ifelse (lean >0, lean,NA),
    rep_lean = ifelse(lean <0, abs(lean), NA)  
  )

ggplot() +
 
  geom_sf(data = MO5_voting, aes(fill = adv_20 / (adv_20 + arv_20)), color = NA) +
  
  scale_fill_gradientn(
    colors = c("#E91D0E", "#FFB6B6", "#B0C4DE", "#0015BC"),
    values = c(0, 0.499, 0.501, 1),
    name = "Partisan Lean"
  ) +
  theme_void() +
  labs(title = "MO Districts 5")
```

...

