---
title: "custom-redist-test"
author: "Fletcher Green"
date: "2026-04-29"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(redist)
library(ggredist)
library(plotly)
source("../BlockManager/Blockmanager.R")
```

# Custom Redistricting Test

## Goal: Make a basic algorithm capable of creating custom district boundaries for two districts based on minority VAP and compactness.

### Section 1: Make Start Districts

```{r warning = F}
nyNewDistricts <- st_read(
  "../shape-files/ny-new-shape/CON24_shapefile_Feb_28_2024/con24.shp"
)

nyNewDistricts <- nyNewDistricts |>
  filter(DISTRICT == 10 | DISTRICT == 11) |>
  select(OBJECTID, DISTRICT, geometry)

nyData2020 <- read.csv("../csv/ny_2020_vtd.csv")
nyData2020 <- nyData2020 |>
  rename(GEOID = GEOID20) |>
  select(pop, nrv, ndv, vap, vap_white, vap_hisp, vap_black, GEOID) |>
  mutate(GEOID = as.character(GEOID))

nyAllBlocks2020 <- BlockManager$GetBlocks("NY")
nyAllBlocks2020 <- BlockManager$JoinByGeoid(nyAllBlocks2020, nyData2020)

nyNewDistricts <- st_make_valid(nyNewDistricts)
centrailizedBlocks <- st_make_valid(nyAllBlocks2020)
centrailizedBlocks <- st_transform(centrailizedBlocks, st_crs(nyNewDistricts))
centrailizedBlocks <- st_point_on_surface(centrailizedBlocks)

distBlocks <- st_join(
  nyNewDistricts,
  centrailizedBlocks,
  join = st_intersects
)
st_geometry(distBlocks) <- st_geometry(nyAllBlocks2020)[
  match(distBlocks$GEOID, nyAllBlocks2020$GEOID)
]

nyWaterBlocks <- NULL
for (i in seq_len(nrow(distBlocks))) {
  if (distBlocks$vap[i] == 0) {
    nyWaterBlocks <- rbind(nyWaterBlocks, distBlocks[i,])
  }
}

nyWaterGeom <- st_union(nyWaterBlocks)

nyWaterGeom <- st_transform(nyWaterGeom, st_crs(nyNewDistricts))
distTenParts <- st_difference(
  nyNewDistricts |> filter(DISTRICT == 10),
  nyWaterGeom
)
distTenParts <- st_cast(distTenParts, "POLYGON")

distElevenParts <- st_difference(
  nyNewDistricts |> filter(DISTRICT == 11),
  nyWaterGeom
)
distElevenParts <- st_cast(distElevenParts, "POLYGON")

largestDistElement <- distElevenParts[1,]
largestDistElementIdx <- 1
secondLargestDistElement <- NULL
secondLargestDistElementIdx <- NULL
for (i in seq_len(nrow(distElevenParts))) {
  if (st_area(distElevenParts[i,]) > st_area(largestDistElement)) {
    secondLargestDistElement <- largestDistElement
    secondLargestDistElementIdx <- largestDistElementIdx
    largestDistElement <- distElevenParts[i,]
    largestDistElementIdx <- i
  }
}

dist11MainlandBlocks <- st_join(
  secondLargestDistElement,
  centrailizedBlocks,
  join = st_intersects
)
st_geometry(dist11MainlandBlocks) <- st_geometry(nyAllBlocks2020)[
  match(dist11MainlandBlocks$GEOID, nyAllBlocks2020$GEOID)
]
dist11MainlandBlocks <- dist11MainlandBlocks |>
      group_by(DISTRICT) |>
      summarize(across(where(is.numeric), sum, na.rm = TRUE))
dist11MainlandPop <- as.numeric(dist11MainlandBlocks$pop[1])
rm(dist11MainlandBlocks)

distElevenParts <- distElevenParts[-secondLargestDistElementIdx,]
distElevenParts <- st_union(distElevenParts, nyWaterGeom)
distElevenParts <- st_snap(distElevenParts, distElevenParts, tolerance = 1)
distElevenParts <- st_union(distElevenParts)
distElevenParts <- st_cast(distElevenParts, "POLYGON")

largestDistElement <- distElevenParts[1,]
for (i in seq_len(length(distElevenParts))) {
  if (st_area(distElevenParts[i,]) > st_area(largestDistElement)) {
    largestDistElement <- distElevenParts[i,]
    largestDistElementIdx <- i
  }
}
distElevenMod <- largestDistElement
distElevenMod <- st_sf(geometry = distElevenMod)
distElevenMod$DISTRICT <- 11
distElevenMod$OBJECTID <- 11


distTenMod <- st_union(
  nyNewDistricts |> filter(DISTRICT == 10 ), 
  secondLargestDistElement
)
distTenMod <- st_difference(
  distTenMod,
  distElevenMod
) |>
  select(DISTRICT, OBJECTID, geometry)

rm(centrailizedBlocks)
rm(distElevenParts)
rm(distBlocks)
rm(distTenParts)
rm(largestDistElement)
rm(nyData2020)
rm(nyWaterGeom)
rm(secondLargestDistElement)
rm(i)
rm(largestDistElementIdx)
rm(secondLargestDistElementIdx)

# Create confirmation plot.
ggplot() +
  geom_sf(data = distTenMod) +
  geom_sf(data = distElevenMod) +
  theme_map()
```

### Section 2: Mainland and Statin Island Redistricting

```{r warning = F}
distElevenTest1 <- distElevenMod
distTenTest1 <- distTenMod
distTenTest1 <- st_cast(distTenTest1, "POLYGON")
distTenTest1 <- distTenTest1[1,]

nyData2020 <- read.csv("../csv/ny_2020_vtd.csv")
nyData2020 <- nyData2020 |>
  rename(GEOID = GEOID20) |>
  select(pop, nrv, ndv, vap, vap_white, vap_hisp, vap_black, GEOID) |>
  mutate(GEOID = as.character(GEOID))

nyAllBlocks2020 <- BlockManager$GetBlocks("NY")
nyAllBlocks2020 <- BlockManager$JoinByGeoid(nyAllBlocks2020, nyData2020)

distTenTest1 <- st_make_valid(distTenTest1)
centrailizedBlocks <- st_make_valid(nyAllBlocks2020)
centrailizedBlocks <- st_transform(centrailizedBlocks, st_crs(distTenTest1))
centrailizedBlocks <- st_point_on_surface(centrailizedBlocks)

distTenBlocks <- st_join(
  distTenTest1,
  centrailizedBlocks,
  join = st_intersects
)
st_geometry(distTenBlocks) <- st_geometry(nyAllBlocks2020)[
  match(distTenBlocks$GEOID, nyAllBlocks2020$GEOID)
]
distTenBlocks <- st_snap(distTenBlocks, distTenBlocks, tolerance = 1)
distTenSumarized <- distTenBlocks |>
      group_by(DISTRICT) |>
      summarize(across(where(is.numeric), sum, na.rm = TRUE))
newDistTenPop <- as.numeric(distTenSumarized$pop[1])
rm(distTenSumarized)

distTenMap <- redist_map(
  distTenBlocks,
  total_pop = pop,
  ndists = ceiling(newDistTenPop / dist11MainlandPop)
)

distTenSim <- redist_smc(distTenMap, nsims = 1, compactness = 0.47)

distTenOptions <- NULL
distElevenData <- NULL
distElevenOptions <- NULL
for (i in 1:ncol(get_plans_matrix(distTenSim))) {
  distTenGroups <- distTenBlocks
  distTenGroups$MAINLANDGROUP <- get_plans_matrix(distTenSim)[, i]
  distTenGroupData <- distTenGroups |>
    group_by(MAINLANDGROUP) |>
    summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
    mutate(pctminority = 1 - (vap_white / vap))
  distTenGroupData <- st_drop_geometry(distTenGroupData)
  distTenGroupGeometry <- distTenGroups |>
    select(MAINLANDGROUP, geometry) |>
    group_by(MAINLANDGROUP) |>
    summarize(geometry = st_union(geometry))
  for (j in 1:ceiling(newDistTenPop / dist11MainlandPop)) {
    continuousDistrict <- st_union(st_geometry(distTenGroupGeometry[-j,]))
    if (!(length(st_cast(continuousDistrict, "POLYGON")) > 1)) {
      distElevenTest1 <- st_transform(distElevenTest1, st_crs(distTenGroupGeometry[j, ]))
      if (any(
        st_touches(
          st_snap(distElevenTest1, distTenGroupGeometry[j,], tolerance = 1),
          distTenGroupGeometry[j, ], 
          sparse = FALSE
        )
      )) {
        distElevenOptions <- bind_rows(distElevenOptions, distTenGroupGeometry[j, ])
        distElevenData <- bind_rows(distElevenData, distTenGroupData[j, ])
        distTenOptions <- bind_rows(
          distTenOptions,
          st_sf(
            st_union(distTenGroupGeometry[-j,])
          )
        )
      }
    }
  }
}

distElevenData <- distElevenData |> arrange(desc(pctminority))
newOrder <- order(distElevenData$pctminority)
distElevenOptions <- distElevenOptions[newOrder,]
distTenOptions <- distTenOptions[newOrder,]
distTenOptions <- distTenOptions |> rename(geometry = st_union.distTenGroupGeometry..j....)
distTenOptions <- st_as_sf(distTenOptions, sf_column_name = "geometry")
distElevenOptions <- distElevenOptions |> select(-MAINLANDGROUP)

rm(centrailizedBlocks)
rm(continuousDistrict)
rm(distElevenData)
rm(distElevenTest1)
rm(distTenBlocks)
rm(distTenGroupData)
rm(distTenGroupGeometry)
rm(distTenGroups)
rm(distTenMap)
rm(distTenSim)
rm(distTenTest1)
rm(nyData2020)
rm(dist11MainlandPop)
rm(i)
rm(j)
rm(newDistTenPop)
rm(newOrder)

# Create confirmation plot. This plot is the district 11 with highest percent minority.
ggplot() +
  geom_sf(data = distTenOptions[1,]) +
  geom_sf(data = distElevenOptions[1,]) +
  theme_map()
```

### Section 3: Spliting Districts 10 and 11 into Two Pieces Redistrictig

```{r warning = F}
distTest <- nyNewDistricts
distTest <- st_cast(distTest, "POLYGON")

nyData2020 <- read.csv("../csv/ny_2020_vtd.csv")
nyData2020 <- nyData2020 |>
  rename(GEOID = GEOID20) |>
  select(pop, nrv, ndv, vap, vap_white, vap_hisp, vap_black, GEOID) |>
  mutate(GEOID = as.character(GEOID))

nyAllBlocks2020 <- BlockManager$GetBlocks("NY")
nyAllBlocks2020 <- BlockManager$JoinByGeoid(nyAllBlocks2020, nyData2020)

distTest <- st_make_valid(distTest)
centrailizedBlocks <- st_make_valid(nyAllBlocks2020)
centrailizedBlocks <- st_transform(centrailizedBlocks, st_crs(distTest))
centrailizedBlocks <- st_point_on_surface(centrailizedBlocks)

testBlocks <- st_join(
  distTest,
  centrailizedBlocks,
  join = st_intersects
)
st_geometry(testBlocks) <- st_geometry(nyAllBlocks2020)[
  match(testBlocks$GEOID, nyAllBlocks2020$GEOID)
]
testBlocks <- st_snap(testBlocks, testBlocks, tolerance = 1)

distMap <- redist_map(
  testBlocks,
  total_pop = pop,
  ndists = 2
)

distSim <- redist_smc(distMap, nsims = 1, compactness = 0.8)

dist1options <- NULL
dist2options <- NULL
dist1data <- NULL
dist2data <- NULL
for (i in 1:ncol(get_plans_matrix(distSim))) {
  distGroups <- testBlocks
  distGroups$MAINLANDGROUP <- get_plans_matrix(distSim)[, i]
  distGroupData <- distGroups |>
    group_by(MAINLANDGROUP) |>
    summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
    mutate(pctminority = 1 - (vap_white / vap))
  distGroupData <- st_drop_geometry(distGroupData)
  distGroupGeometry <- distGroups |>
    select(MAINLANDGROUP, geometry) |>
    group_by(MAINLANDGROUP) |>
    summarize(geometry = st_union(geometry))
  
  dist1data <- bind_rows(dist1data, distGroupData |> filter(MAINLANDGROUP == 1))
  dist2data <- bind_rows(dist2data, distGroupData |> filter(MAINLANDGROUP == 2))
  dist1options <- bind_rows(dist1options, distGroupGeometry |> filter(MAINLANDGROUP == 1))
  dist2options <- bind_rows(dist2options, distGroupGeometry |> filter(MAINLANDGROUP == 2))
}

dist1data <- dist1data |> arrange(desc(pctminority))
newOrder <- order(dist1data$pctminority)
dist1options <- dist1options[newOrder,]
dist2options <- dist2options[newOrder,]

rm(centrailizedBlocks)
rm(dist1data)
rm(dist2data)
rm(distGroupData)
rm(distGroupGeometry)
rm(distMap)
rm(distSim)
rm(distTen)
rm(distTest)
rm(nyData2020)
rm(testBlocks)
rm(distGroups)
rm(i)
rm(newOrder)

# Create confirmation plot. This plot is the district 11 with highest percent minority.
ggplot() +
  geom_sf(data = dist1options[1,]) +
  geom_sf(data = dist2options[1,]) +
  theme_map()
```

### Section 4: Plotting 2 Parts Redistricting

```{r warning = F}
  d1 <- dist1options[1,] |>
    select(geometry) |>
    mutate(OBJECTID = 10) |>
    mutate(DISTRICT = 10)

  d2 <- dist2options[1,] |>
    select(geometry) |>
    mutate(OBJECTID = 11) |>
    mutate(DISTRICT = 11)
  distElevenPlot <- BlockManager$ConsolodateDistricts(d1, nyAllBlocks2020)
  distTenPlot <- BlockManager$ConsolodateDistricts(d2, nyAllBlocks2020)

  ggplot() +
    geom_sf(data = distTenPlot, aes(fill = ndv / (nrv + ndv))) +
    geom_sf(data = distElevenPlot, aes(fill = ndv / (nrv + ndv))) +
    geom_sf(data = nyWaterBlocks, fill = "gray", color = NA) +
    scale_fill_party_c() +
    theme_map()
```

### Section 6: Plotting Mainland + Statin Island Redistricting

```{r warning = F}
PlotVoteShare <- function(ten, eleven) {
  distElevenPlot <- st_union(
    distElevenMod, 
    st_transform(eleven, crs = st_crs(distElevenMod))
  )
  distTenPlot <- ten
  distTenPlot <- distTenPlot |>
    mutate(DISTRICT = 10) |>
    mutate(OBJECTID = 10)

  distElevenPlot <- BlockManager$ConsolodateDistricts(distElevenPlot, nyAllBlocks2020)
  distTenPlot <- BlockManager$ConsolodateDistricts(distTenPlot, nyAllBlocks2020)

  ggplot() +
    geom_sf(data = distTenPlot, aes(fill = ndv / (nrv + ndv))) +
    geom_sf(data = distElevenPlot, aes(fill = ndv / (nrv + ndv))) +
    geom_sf(data = nyWaterBlocks, fill = "gray", color = NA) +
    scale_fill_party_c() +
    theme_map()
}

PlotMinority <- function(ten, eleven) {
  distElevenPlot <- st_union(
    distElevenMod, 
    st_transform(eleven, crs = st_crs(distElevenMod))
  )
  distTenPlot <- ten
  distTenPlot <- distTenPlot |>
    mutate(DISTRICT = 10) |>
    mutate(OBJECTID = 10)

  distElevenPlot <- BlockManager$ConsolodateDistricts(distElevenPlot, nyAllBlocks2020)
  distTenPlot <- BlockManager$ConsolodateDistricts(distTenPlot, nyAllBlocks2020)

  ggplot() +
    geom_sf(data = distTenPlot, aes(fill = 1 - (vap_white / vap))) +
    geom_sf(data = distElevenPlot, aes(fill = 1 - (vap_white / vap))) +
    geom_sf(data = nyWaterBlocks, fill = "gray", color = NA) +
    scale_fill_gradient2(
      low = "#461055",
      mid = "#2A9A86",
      high = "#FAEC5E",
      midpoint = 0.5,
      limits = c(0, 1)
    ) +
    theme_map() +
    labs(
      fill = "Percent Minority"
    )
}

nyNewDistricts <- st_read(
  "../shape-files/ny-new-shape/CON24_shapefile_Feb_28_2024/con24.shp"
)

dist10 <- nyNewDistricts |>
  filter(DISTRICT == 10) |>
  select(OBJECTID, DISTRICT, geometry)
dist11 <- nyNewDistricts |>
  filter(DISTRICT == 11) |>
  select(OBJECTID, DISTRICT, geometry)

PlotVoteShare(distTenOptions[1,], distElevenOptions[1,])
PlotMinority(distTenOptions[1,], distElevenOptions[1,])

PlotVoteShare(dist10, dist11)
PlotMinority(dist10, dist11)
```

### Section 7: Saving Mainland + Statin Island Redistricting

```{r warning = F, eval = F}
distTenOptions <- distTenOptions |>
  mutate(DISTRICT = 10) |>
  mutate(OBJECTID = 10)

distElevenOptions <- distElevenOptions |>
  mutate(DISTRICT = 11) |>
  mutate(OBJECTID = 11)


dist11_union <- st_union(
  st_transform(distElevenOptions[1,], st_crs(distElevenMod)),
  distElevenMod
)
dist11_union <- st_sf(geometry = dist11_union) |>
  mutate(DISTRICT = 11) |>
  mutate(OBJECTID = 11) |>
  select(geometry, DISTRICT, OBJECTID)
ggplot() +
  geom_sf(data = dist11_union)

mergedDistricts <- rbind(
  st_transform(distTenOptions[1,], crs = st_crs(distElevenMod)),
  dist11_union
)

st_write(mergedDistricts, "./shp/custom-bounds.shp")
```