---
title: "District 11 Exploration"
author: "Fletcher Green"
date: "2026-04-14"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vistime)
library(redist)
library(ggredist)
source("../BlockManager/Blockmanager.R")
```

# New York Redistricting Timeline

- **Description:**
  - A timeline of redistricting events in New York following the 2020 census. Data manually collected from: <https://redistricting.lls.edu/state/new-york/?cycle=2020&level=Congress&startdate=2022-02-03>.

```{r warning=F, message=F, echo=F}
districtData <- data.frame(
  group = c(
    rep(c("Approved\nPlan\nDurations"), each = 4),
    rep(c("Group\nin\nControl"), each = 6)
  ),
  names = c(
    "", "L", "C", "L",
    "R", "L", "C", "R", "L", "R"
  ),
  start = c(
    "2020-01-01", "2022-02-03", "2022-06-02", "2024-02-28",
    "2020-01-01", "2022-01-03", "2022-04-27", "2022-06-02", "2024-02-15", "2024-02-28"
  ),
  end = c(
    "2022-02-03", "2022-06-02", "2024-02-28", "2026-04-14",
    "2022-01-03", "2022-04-27", "2022-06-02", "2024-02-15", "2024-02-28", "2026-04-14"
  ),
  color = c(
    "gray", "orange", "red", "orange", 
    "blue", "orange", "red", "blue", "orange", "blue"
  )
)

gg_vistime(districtData, col.event = "names", title = "New York Redistricting Timeline")
```

# New York Extended Minority Map

- **Description:**
  - An extended version of Nick's minority map which attempts to better capture what is happening in district 11.

```{r warning=F, message=F, echo=F}
# Read geometry of districts.
nyNewDistricts <- st_read("../shape-files/ny-new-shape/CON24_shapefile_Feb_28_2024/con24.shp")

# Remove unnecessary data and rename cols so that they match what is expected
# by BlockManager.
nyNewDistrictBlocks <- nyNewDistricts |>
  select(OBJECTID, DISTRICT, geometry)

dist11Adj <- st_touches(nyNewDistrictBlocks)[11]

nyNewDistrictBoundaries <- nyNewDistrictBlocks[c(11, dist11Adj[[1]]),]
nyNewDistrictBlocks <- nyNewDistrictBlocks[c(11, dist11Adj[[1]]),]

# Read block level data. Change col names so that they match what is
# expected by BlockManager.
nyData2020 <- read.csv("../csv/ny_2020_vtd.csv")

nyData2020 <- nyData2020 |>
  rename(GEOID = GEOID20) |>
  select(nrv, ndv, vap, vap_white, GEOID) |>
  mutate(GEOID = as.character(GEOID))

# Get blocks and join block level data to it.
nyAllBlocks2020 <- BlockManager$GetBlocks("NY")
nyAllBlocks2020 <- BlockManager$JoinByGeoid(nyAllBlocks2020, nyData2020)

# Translate CRS and make blocks into centroids.
nyNewDistrictBlocks <- st_make_valid(nyNewDistrictBlocks)
centrailizedBlocks <- st_make_valid(nyAllBlocks2020)
centrailizedBlocks <- st_transform(centrailizedBlocks, st_crs(nyNewDistrictBlocks))
centrailizedBlocks <- st_point_on_surface(centrailizedBlocks)

# Perform spatial join.
nyNewDistrictBlocks <- st_join(
  nyNewDistrictBlocks,
  centrailizedBlocks,
  join = st_intersects
)

st_geometry(nyNewDistrictBlocks) <- st_geometry(nyAllBlocks2020)[
  match(nyNewDistrictBlocks$GEOID, nyAllBlocks2020$GEOID)
]

nyWaterBlocks <- NULL
for (i in seq_len(nrow(nyNewDistrictBlocks))) {
  if (nyNewDistrictBlocks$vap[i] == 0) {
    nyWaterBlocks <- rbind(nyWaterBlocks, nyNewDistrictBlocks[i,])
  }
}

# Plot districts.
ggplot() +
  geom_sf(
    data = nyNewDistrictBlocks, 
    color = NA, 
    aes(fill = 1 - (vap_white / vap)), 
    linewidth = 0.5
  ) +
  scale_fill_gradient2(
    low = "#461055",
    mid = "#2A9A86",
    high = "#FAEC5E",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_map() +
  geom_sf(
    data = nyWaterBlocks, 
    fill = adjustcolor("blue", alpha.f = 0.5), 
    color = NA, 
    linewidth = 1.5) +
  geom_sf(
    data = nyNewDistrictBoundaries |> filter(DISTRICT == 10), 
    color = "blue", 
    linewidth = 2.0
  ) +
  geom_sf(
    data = nyNewDistrictBoundaries |> filter(DISTRICT == 11), 
    color = "red", 
    linewidth = 2.0
  ) +
  geom_sf(
    data = nyNewDistrictBoundaries, 
    color = adjustcolor("#2B2B2B", alpha.f = 0.9), 
    linewidth = 1.0
  ) +
  geom_sf_text(
    data = nyNewDistrictBoundaries,
    aes(label = DISTRICT),
    size = 8,
    color = "white"
  )
```

# New York Extended Vote Share Map

- **Description:**
  - Recolored version of the minority map above attempting to display vote share.

```{r warning=F, message=F, echo=F}
# Read geometry of districts.
nyNewDistricts <- st_read("../shape-files/ny-new-shape/CON24_shapefile_Feb_28_2024/con24.shp")

# Remove unnecessary data and rename cols so that they match what is expected
# by BlockManager.
nyNewDistrictBlocks <- nyNewDistricts |>
  select(OBJECTID, DISTRICT, geometry)

dist11Adj <- st_touches(nyNewDistrictBlocks)[11]

nyNewDistrictBoundaries <- nyNewDistrictBlocks[c(11, dist11Adj[[1]]),]
nyNewDistrictBlocks <- nyNewDistrictBlocks[c(11, dist11Adj[[1]]),]

# Read block level data. Change col names so that they match what is
# expected by BlockManager.
nyData2020 <- read.csv("../csv/ny_2020_vtd.csv")

nyData2020 <- nyData2020 |>
  rename(GEOID = GEOID20) |>
  select(nrv, ndv, vap, vap_white, GEOID) |>
  mutate(GEOID = as.character(GEOID))

# Get blocks and join block level data to it.
nyAllBlocks2020 <- BlockManager$GetBlocks("NY")
nyAllBlocks2020 <- BlockManager$JoinByGeoid(nyAllBlocks2020, nyData2020)

# Translate CRS and make blocks into centroids.
nyNewDistrictBlocks <- st_make_valid(nyNewDistrictBlocks)
centrailizedBlocks <- st_make_valid(nyAllBlocks2020)
centrailizedBlocks <- st_transform(centrailizedBlocks, st_crs(nyNewDistrictBlocks))
centrailizedBlocks <- st_point_on_surface(centrailizedBlocks)

# Perform spatial join.
nyNewDistrictBlocks <- st_join(
  nyNewDistrictBlocks,
  centrailizedBlocks,
  join = st_intersects
)

st_geometry(nyNewDistrictBlocks) <- st_geometry(nyAllBlocks2020)[
  match(nyNewDistrictBlocks$GEOID, nyAllBlocks2020$GEOID)
]

nyWaterBlocks <- NULL
for (i in seq_len(nrow(nyNewDistrictBlocks))) {
  if (nyNewDistrictBlocks$vap[i] == 0) {
    nyWaterBlocks <- rbind(nyWaterBlocks, nyNewDistrictBlocks[i,])
  }
}

# Plot districts.
ggplot() +
  geom_sf(
    data = nyNewDistrictBlocks, 
    color = NA, 
    aes(fill = ndv / (ndv + nrv)), 
    linewidth = 0.5
  ) +
  scale_fill_party_c() +
  theme_map() +
  geom_sf(
    data = nyWaterBlocks, 
    fill = adjustcolor("gray", alpha.f = 1.0), 
    color = NA, 
    linewidth = 1.5) +
  geom_sf(
    data = nyNewDistrictBoundaries |> filter(DISTRICT == 10), 
    color = "blue", 
    linewidth = 2.0
  ) +
  geom_sf(
    data = nyNewDistrictBoundaries |> filter(DISTRICT == 11), 
    color = "red", 
    linewidth = 2.0
  ) +
  geom_sf(
    data = nyNewDistrictBoundaries, 
    color = adjustcolor("#2B2B2B", alpha.f = 0.9), 
    linewidth = 1.0
  ) +
  geom_sf_text(
    data = nyNewDistrictBoundaries,
    aes(label = DISTRICT),
    size = 8,
    color = "white"
  )
```