---
title: "District Comparison"
author: "Fletcher Green"
date: "2026-04-01"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(alarmdata)
library(redist)
library(ggredist)
source("BlockManager/Blockmanager.R")
library(tidyverse)
library(ggredist)
library(sf)
```

# District Comparison

- **Overview:** Both New York and Missouri have redistricted twice since the 2020 census. New York is undergoing litigation and may redistrict a third time. This document compares the first two redistrictings in each state.

## Missouri

- **Version:** This is the first map enacted by Missouri following the 2020 census.
- **Link to data:** <https://redistrictingdatahub.org/dataset/2022-missouri-congressional-districts-approved-plan/>

```{r warning=F, message=F, echo=F}
mapMoOld <- st_read("shape-files/mo-init-shape/5799H_02T.shp")
mapMoOld |> 
  ggplot() +
  geom_sf() +
  theme_map()
```

- **Version:** This is the second map enacted by Missouri following the 2020 census.
- **Link to data:** <https://redistrictingdatahub.org/dataset/2025-missouri-congressional-districts-plan/>

```{r warning=F, echo=F}
mapMoNew <- st_read("shape-files/mo-new-shape/HB1_Missouri_Congressional_Districts_2025.shp")
mapMoNew |> 
  ggplot() +
  geom_sf() +
  theme_map()
```

## New York

- **Version:** This is the first map enacted by New York following the 2020 census.
- **Link to data:** <https://redistrictingdatahub.org/dataset/2022-new-york-congressional-districts-plan/>

```{r warning=F, message=F, echo=F}
mapNyOld <- st_read("shape-files/ny-init-shape/CON22_June_03_2022.shp")
mapNyOld |> 
  ggplot() +
  geom_sf() +
  theme_map()
```

- **Version:** This is the second map enacted by New York following the 2020 census.
- **Link to data:** <https://redistrictingdatahub.org/dataset/2024-new-york-congressional-districts-plan-approved/>

```{r warning=F, echo=F}
mapNyNew <- st_read("shape-files/ny-new-shape/CON24_shapefile_Feb_28_2024/con24.shp")
mapNyNew |> 
  ggplot() +
  geom_sf() +
  theme_map()
```

# Improved Plots

```{r warning=F, echo=F}
moNewDistricts <- st_read(
  "shape-files/mo-new-shape/HB1_Missouri_Congressional_Districts_2025.shp"
)
moNewDistricts <- moNewDistricts |>
  select(OBJECTID, District, geometry) |>
  rename(DISTRICT = District)

moData2020 <- read.csv("mo_2020_vtd.csv")
moData2020 <- moData2020 |>
  rename(GEOID = GEOID20) |>
  select(nrv, ndv, GEOID) |>
  mutate(GEOID = as.character(GEOID))

moBlocks2020 <- BlockManager$GetBlocks("MO")
moBlocks2020 <- BlockManager$JoinByGeoid(moBlocks2020, moData2020)

moNewDistricts <- BlockManager$ConsolodateDistricts(moNewDistricts, moBlocks2020)

ggplot() +
  geom_sf(data = moNewDistricts, aes(fill = ndv / (ndv + nrv)), linewidth = 0.7) +
  scale_fill_party_c() +
  theme_map() +
  geom_district(data = moNewDistricts, fill = NA, aes(group = TRUE), linewidth = 1.0)
```

```{r warning=F, echo=F}
moOldDistricts <- st_read("shape-files/mo-init-shape/5799H_02T.shp")
moOldDistricts <- moOldDistricts |>
  select(ID, DISTRICT, geometry) |>
  rename(OBJECTID = ID) |>
  mutate(DISTRICT = as.numeric(DISTRICT))

moOldDistricts <- BlockManager$ConsolodateDistricts(moOldDistricts, moBlocks2020)

ggplot() +
  geom_sf(data = moOldDistricts, aes(fill = ndv / (ndv + nrv)), linewidth = 0.7) +
  scale_fill_party_c() +
  theme_map() +
  geom_district(data = moNewDistricts, fill = NA, aes(group = TRUE), linewidth = 1.0)
```

```{r warning=F, echo=F}
nyNewDistricts <- st_read("shape-files/ny-new-shape/CON24_shapefile_Feb_28_2024/con24.shp")

nyNewDistricts <- nyNewDistricts |>
  select(OBJECTID, DISTRICT, geometry)

nyData2020 <- read.csv("ny_2020_vtd.csv")
nyData2020 <- nyData2020 |>
  rename(GEOID = GEOID20) |>
  select(nrv, ndv, GEOID) |>
  mutate(GEOID = as.character(GEOID))

nyBlocks2020 <- BlockManager$GetBlocks("NY")
nyBlocks2020 <- BlockManager$JoinByGeoid(nyBlocks2020, nyData2020)

nyNewDistricts <- BlockManager$ConsolodateDistricts(nyNewDistricts, nyBlocks2020)
sum(nyNewDistricts$nrv)
ggplot() +
  geom_sf(data = nyNewDistricts, aes(fill = ndv / (ndv + nrv)), linewidth = 0.7) +
  scale_fill_party_c() +
  theme_map() +
  geom_district(data = nyNewDistricts, fill = NA, aes(group = TRUE), linewidth = 1.0)
```

```{r warning=F, echo=F}
nyOldDistricts <- st_read("shape-files/ny-init-shape/CON22_June_03_2022.shp")

nyOldDistricts <- mapNyOld |>
  select(OBJECTID, DISTRICT, geometry)

nyOldDistricts <- BlockManager$ConsolodateDistricts(nyOldDistricts, nyBlocks2020)

ggplot() +
  geom_sf(data = nyOldDistricts, aes(fill = ndv / (ndv + nrv)), linewidth = 0.7) +
  scale_fill_party_c() +
  theme_map() +
  geom_district(data = nyNewDistricts, fill = NA, aes(group = TRUE), linewidth = 1.0)
```

```{r warning=F, echo=F}
mapMoNew <- st_read("shape-files/mo-new-shape/HB1_Missouri_Congressional_Districts_2025.shp")
mapMoOld <- st_read("shape-files/mo-init-shape/5799H_02T.shp")
mapMoNew <- mapMoNew |>
  select(OBJECTID, District, geometry) |>
  rename(DISTRICT = District) |>
  st_transform(st_crs(mapMoOld))

mapMoOld <- mapMoOld |>
  select(ID, DISTRICT, geometry) |>
  rename(OBJECTID = ID)

mapMoOverlap <- NULL
for (i in seq_len(nrow(mapMoNew))) {
  mapMoOverlap <- rbind(mapMoOverlap, st_intersection(mapMoNew[i,], mapMoOld[i,]))
}

ggplot() +
  geom_district(
    data = mapMoNew, 
    color = NA, 
    fill = "white", 
    aes(group = TRUE), 
    linewidth = 1.5
  ) +
  geom_sf(data = mapMoOverlap, fill = "orange", color = NA, linewidth = 0.7) +
  geom_sf(data = mapMoNew, fill = NA, linewidth = 0.7) +
  geom_district(data = mapMoNew, fill = NA, aes(group = TRUE), linewidth = 1.0) +
  geom_sf(
    data = mapMoOld[5,],
    color = "blue",
    fill = adjustcolor("#1F77B4", alpha.f = 0.3),
    linewidth = 1.2
  ) +
  theme_map()
```