---
title: "District Comparison"
author: "Fletcher Green"
date: "2026-04-01"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggredist)
library(sf)
```

# District Comparison

- **Overview:** Both New York and Missouri have redistricted twice since the 2020 census. New York is undergoing litigation and may redistrict a third time. This document compares the first two redistrictings in each state.

## Missouri

- **Version:** This is the first map enacted by Missouri following the 2020 census.
- **Link to data:** <https://redistrictingdatahub.org/dataset/2022-missouri-congressional-districts-approved-plan/>

```{r warning=F, message=F, echo=F}
mapMoOld <- st_read("shape-files/mo-init-shape/5799H_02T.shp")
mapMoOld |> 
  ggplot() +
  geom_sf() +
  theme_map()
```

- **Version:** This is the second map enacted by Missouri following the 2020 census.
- **Link to data:** <https://redistrictingdatahub.org/dataset/2025-missouri-congressional-districts-plan/>

```{r warning=F, echo=F}
mapMoNew <- st_read("shape-files/mo-new-shape/HB1_Missouri_Congressional_Districts_2025.shp")
mapMoNew |> 
  ggplot() +
  geom_sf() +
  theme_map()
```

## New York

- **Version:** This is the first map enacted by New York following the 2020 census.
- **Link to data:** <https://redistrictingdatahub.org/dataset/2022-new-york-congressional-districts-plan/>

```{r warning=F, message=F, echo=F}
mapNyOld <- st_read("shape-files/ny-init-shape/CON22_June_03_2022.shp")
mapNyOld |> 
  ggplot() +
  geom_sf() +
  theme_map()
```

- **Version:** This is the second map enacted by New York following the 2020 census.
- **Link to data:** <https://redistrictingdatahub.org/dataset/2024-new-york-congressional-districts-plan-approved/>

```{r warning=F, echo=F}
mapNyNew <- st_read("shape-files/ny-new-shape/CON24_shapefile_Feb_28_2024/con24.shp")
mapNyNew |> 
  ggplot() +
  geom_sf() +
  theme_map()
```
